SOMETHING ABOUT THE DISTRIBUTED AVERAGING BASED APPROXIMATION IN SENSOR NETWORKS

### 2022-10-11 paper.tex
tuesday party

\section{DISTRIBUTED ADAPTIVE BLIND SYSTEM IDENTIFICATION USING ONLINE-ADMM}
\label{sec:dbsi}
In \cite{blochbergerDBSI}, we proposed an adaptive SIMO BSI algorithm based on Online-ADMM.
In this section, we will introduce parts which are necessary for the reader to understand the extension.
For the full derivation and algorithm, we refer the reader to the cited publication.

The coss-relation problem formulation is used relative information between sensor signals to identify the system, i.e., the acoutic or communication channels.
The solution to this problem is found by the minimization problem
\begin{equation}
  \begin{aligned}
      \hat{\h} = \arg \min_{\h} \quad &\h^\herm \hat{\R} \h \\
      \text{s.t. } \quad &\h^\herm \h = 1.
  \end{aligned}\label{eq:frequency_domain:min_prob}
\end{equation}
where 
\begin{align}
  \w_i^{k+1} &= \underset{\w_i}{\operatorname{argmin}} \, \mathcal{L}_{\rho} (\w,\h^k,\uu^k)\label{eq:general_consensus_admm:local}\\
  \h^{k+1} &= \underset{\h, \|\h\| = 1}{\operatorname{argmin}}\, \mathcal{L}_{\rho} (\w^{k+1},\h,\uu^k)\label{eq:general_consensus_admm:global}\\
  \uu_i^{k+1} &= \uu_i^{k} + \rho \left( \w_i^{k+1} - \tilde{\h}_i^{k+1} \right),\label{eq:general_consensus_admm:dual}
\end{align}

% The minimization problem for the local variable \(\w_i\) \eqref{eq:general_consensus_admm:local} can be solved by various algorithms, in this case however we perform the adaptive update step as
% \begin{equation}
%     \w_i^{m+1} = \w_i^{m} - \mu \bm{{V}}_i^m \left( \hat{\aRhof}_i^m \w_i^m + \uu_i^m + \rho\left(\w_i^m - \tilde{\h}_i^{m}\right)\right),\label{eq:online_admm:local_update}
% \end{equation}
% where \(\mu\), (\(0  < \mu\leq 1\)), is a step size and \(\bm{{V}}_i^m = \left(\hat{\aRhof}_i^m + \rho \I \right)^{-1}\) is the regularized inverse Hessian of the problem.
% The implicit overlap-save matrices in \(\bm{{V}}_i^m\) make this a constrained update step which is costly to compute, so in order to reduce computational complexity, we introduce the approximation
% \begin{equation}
%     \hat{\bm{{V}}}_i^m = \operatorname{diag} \left\{ \operatorname{diag} \left\{ \hat{\aRhof}_i^m \right\} + \rho \mtxb{1} \right\}^{-1},
% \end{equation}
% which is straightforward to compute.
It may be readily verified (cf. \cite{boydDistributedOptimizationStatistical2011}) that the solution to \eqref{eq:general_consensus_admm:global} is given by an update step which computes
\begin{equation}
    \h_i^{m+1} = \frac{\bar{\h}_i^{m+1}}{\sqrt{\sum_{j \in \Mset} \|\bar{\h}_j^{m+1}\|^2}}\label{eq:online_admm:consensus_update}
\end{equation}
for each node \(i \in \Mset\) with the local unnormalized consensus
\begin{equation}
    \bar{\h}_i^{m+1} = \bar{\w}_i^{m+1} + \frac{1}{\rho} \bar{\uu}_i^{m}.
\end{equation}
Here, \(\bar{\w}_i = \frac{1}{N_i} \sum_{j \in \Tset_i} (\w_j)_i\) and \(\bar{\uu}_i = \frac{1}{N_i} \sum_{j \in \Tset_i} (\uu_j)_i\), with \(N_i = | \Tset_i |\), are neigborhood averages of the channel response estimate and the dual variable of channel \(i\) respectively, computed at node \(i\). The values of all \(\|\bar{\h}_i^{m+1}\|^2\), \(i \in \Mset\) are transmitted and relayed through the network until all nodes can compute the denominator in \eqref{eq:online_admm:consensus_update}.
% Finally, the update of the dual variables is given by
% \begin{equation}
%     \uu_i^{m+1} = \uu_i^{m} + \rho \left( \w_i^{m+1} - \tilde{\h}_i^{m+1} \right).\label{eq:online_admm:dual_update}
% \end{equation}

POINT: update \eqref{eq:online_admm:consensus_update} requires knowledge of all \(\|\bar{\h}_i^{m+1}\|^2\), \(i \in \Mset\) and we introduce a distributed averaging-based approach to tackle this.

\section{Adaptive Norm Estimation}
\label{sec:adaptivenormest}
Distributed averaging:
\begin{mini}{}{\|\W - \bm{1}\bm{1}^\T/n\|_2} {\label{eq1}}{}
  \addConstraint{\W}{\in \mathcal{C}}
  \addConstraint{\bm{1}^\T \W}{= \bm{1}^\T}
  \addConstraint{\W \bm{1}}{= \bm{1}}
\end{mini}
gives us the optimal weights \(\mtxb{W}_{o}\) where the weights used by node \(i\) to combine values from neighbor nodes \(j \in \Rset_i\) are given at the \(i\)-th row. The distributed averaging approach is an iterative approach, however here, we will try to find a way to limit the number of iterations necessary. This goal comes from the fact that the iterations require data from neighboring nodes leading to additional transmissions (per frame).

% We take avantage of the fact that the squared norm of the stacked solution vector can be separated into
% \begin{equation}
%   \| \h \|^2 = \sum_{j \in \Mset} \|\h_j\|^2.
% \end{equation}
We define the approximation of the global norm \(\hat{\phi}_i \approx \|\h\|^2\) at each node \(i\) at frame \(m\) as follows:
\begin{equation}
  \phi_i^{m,k+1} = \sum_{j \in \Rset_i} w_{ij} \phi_j^{m,k}\quad\text{for }k=1,\ldots,K\label{eq:dist_norm_est_iter}
\end{equation}
with the initial value at \(m=1,k=1\) defined as \(\phi_i^{1,1}=\|\h_i^{1}\|^2\).
\eqref{eq:dist_norm_est_iter} is applied \(K\) times.
We now define the frame estimate of the global squared norm as
\begin{equation}
  \hat{\phi}_i^{m+1} = \gamma_i \bar{\phi}_i^{m} + (1-\gamma) \phi_i^{m,K}
\end{equation}
where \(\bar{\phi}_i^{m} = \sum_{j \in \Rset_i} w_{ij} \|\h_i^m\|^2\) is an instantaneous weighted neighborhood average of node-wise squared norm values and \(\gamma_i\) is a mixing factor.
This factor can be chosen as a fixed value, where simulations have shown that values around \(\gamma_i = 0.02\) lead to satisfying results (cf. FIGURE) or adaptively.
For this, we use the instantaneous values to track stationarity of the squared norms by setting the factor \(\gamma_i^{m}\) proportional to \(| \bar{\phi}_i^{m} - \bar{\phi}_i^{m-1} |\), the absolute difference between instantaneous neighborhood norm values of subsequent frames.

The special case of \(K=1\) leads to the update step:
\begin{equation}
  \hat{\phi}_i^{m+1} = \gamma_i \sum_{j \in \Rset_i} w_{ij} \| \bar{\h}_j^{m+1} \|^2 + (1-\gamma_i) \sum_{j \in \Rset_i} w_{ij} \hat{\phi}_j^m
\end{equation}

@@

% \begin{align}
%   \w_i^{m+1} &= \operatorname{argmin} \mathcal{L}_\rho(\w,\h,\uu)\\
%   \bar{\h}_i^{m+1} &= \bar{\w}_i^{m+1} + \frac{1}{\rho} \bar{\uu}_i^m\\
%   \phi_i^1 &= \hat{\phi}_i^{m}\\
%   \phi_i^{k+1} &= \sum_{j \in \Rset_i} w_{ij} \phi_j^{k} \qquad\text{for } k=1,\ldots,K\\
%   \hat{\phi}_i^{m+1} &= \gamma \| \bar{\h}_j^{m+1} \|^2 + (1-\gamma) \phi_i^K\\
%   \h_i^{m+1} &= \frac{\bar{\h}_i^{m+1}}{\sqrt{\hat{\phi}_i^K\,M}}\\
%   \uu_i^{m+1} &= \uu_i^{m} + \rho \left(\w_i^{m+1} - \tilde{\h}_i^{m+1}\right).
% \end{align}

>>>>

\section{DISTRIBUTED ADAPTIVE BSI IN SENSOR NETWORKS WITH ONLINE-ADMM}
\label{sec:dbsi}
\begin{todo}
  Introduce signal model, \(\h\),...
\end{todo}
% In \cite{blochbergerDBSI}, we proposed an adaptive SIMO BSI algorithm based on Online-ADMM.
In this section, we will introduce certain parts of an adaptive SIMO BSI algorithm based on Online-ADMM \cite{blochbergerDBSI}, which are necessary for the reader to understand its extension of it.
We refer the reader to the publication cited above for the full derivation of the algorithm.

We assume a set \(\Mset \triangleq \{1,\ldots,M\}\) sensor nodes and a set of edges \(\mathcal{E}\) which connect the nodes forming a sensor network.
Each edge is an unordered pair of nodes \(\{i,j\} \in \mathcal{E}\), which represents the communication link between them.
From this follows the neighborhood of a node \(i\), i.e., the other nodes it is directly connected to, defined as \(\Nset_i = \{j|\{i,j\} \in \mathcal{E}\}\).
Furthermore, we can define the symmetric adjacency matrix \(\mtxb{C}\) with elements
\begin{equation}
  c_{ij} = \begin{cases}
    1 \quad \text{if } i = j\\
    1 \quad \text{if } i \neq j \text{ and }\{i,j\} \in \mathcal{E}\\
    0 \quad \text{otherwise}.
  \end{cases}
\end{equation}
It is to note that in \cite{blochbergerDBSI}, the pairs \(\{i,j\} \in \mathcal{E}\) are ordered, which yields a directed graph and leads to two sets of neighborhoods ("transmit" and "receive") and therefore a non-symmetric adjacency matrix.
In this paper, we limit the explanation to the undirected graph to save space, as the extension is straightforward.

In BSI, the coss-relation problem formulation uses relative information between sensor signals to identify the system, i.e., the acoustic or communication channels.
The solution to this problem is found by the minimization problem (cf. e.g. \cite{langtongBlindIdentificationEqualization1994,huangAdaptiveMultichannelLeast2002,huangClassFrequencydomainAdaptive2003})
\begin{equation}
  \begin{aligned}
      \hat{\h} = \arg \min_{\h} \quad &\h^\herm \hat{\R} \h \\
      \text{s.t. } \quad &\h^\herm \h = 1.
  \end{aligned}\label{eq:frequency_domain:min_prob}
\end{equation}
\begin{todo}
  Introduce \(\w, \uu\)
\end{todo}
Separating the cost function and appying general-form consensus ADMM as an adaptive algorithm leads to the update steps
\begin{align}
  \w_i^{m+1} &= \underset{\w_i}{\operatorname{argmin}} \, \mathcal{L}_{\rho} (\w,\h^m,\uu^m)\label{eq:general_consensus_admm:local}\\
  \h^{m+1} &= \underset{\h, \|\h\| = 1}{\operatorname{argmin}}\, \mathcal{L}_{\rho} (\w^{m+1},\h,\uu^m)\label{eq:general_consensus_admm:global}\\
  \uu_i^{m+1} &= \uu_i^{m} + \rho \left( \w_i^{m+1} - \tilde{\h}_i^{m+1} \right).\label{eq:general_consensus_admm:dual}
\end{align}
Here, the second step, the update of the so-called consensus variables are of interest for this extension.
It is defined \cite{blochbergerDBSI} as
\begin{equation}
    \h_i^{m+1} = \frac{\bar{\h}_i^{m+1}}{\|\h^{m+1}\|}\label{eq:online_admm:consensus_update}
\end{equation}
for each node \(i \in \Mset\) with the local unnormalized consensus \(\bar{\h}_i^{m+1}\), a combination of neigborhood averages of the channel response estimates and the dual variables of channel \(i\) respectively, computed at node \(i\).
The denominator of \eqref{eq:online_admm:consensus_update}, \(\|\h\|\), is the stacked vector of the full system. This vector, however, is not available at any of the nodes in the network.
In \cite{blochbergerDBSI}, we assume that partial squared norms of \(\|\h\|^2 = \sum_{i \in \Mset} \|\h_i\|^2\) are relayed throughout the network until all nodes have the information necessary.
In \cite{yuDistributedBlindSystem2014,liuDistributedBlindIdentification2016}, a similar global value is estimated by a distributed averaging approach, iteratively combining values within neighborhoods until convergence to a network-wide average.

Subsequently, we introduce an extension to the algorithm described in \cite{blochbergerDBSI} using a fastest distributed linear averaging (FDLA) approach \cite{xiaoFastLinearIterations2004} with the further introduction of a mixing factor to include instantaneous values in the recursion.

\section{FDLA-based adaptive norm estimation}
\label{sec:adaptivenormest}
We define 
The FDLA problem, as introduced in \cite{xiaoFastLinearIterations2004}, looks to find optimal weights for distributed linear combinations of the form \(\bm{\phi}^{k+1} = \W \bm{\phi}^{k}\) where \(\bm{\phi}^k = \begin{bmatrix} \phi_1^k & \ldots & \phi_M^k \end{bmatrix}^\T\) is a vector, which initially, at \(k=1\), contains a set of variables, of which an average should be computed iteratively.
The problem comes in the form of a minimization problem
\begin{equation}
  \begin{aligned}
    \min& \quad \| \W - \bm{1}\bm{1}^\T/n\|_2\\
    \text{s.t.}& \quad \W \in \mathcal{C},\, \bm{1}^\T \W = \bm{1}^\T,\, \W \bm{1}= \bm{1}
  \end{aligned}
\end{equation}

The set \(\mathcal{C} = \{\W \in \mathbb{R}^{M \times M} \vert W_{ij} = 0 \text{ if } \{i,j\} \notin \mathcal{E}\}\) describes all matrices with the same sparsity pattern as the adjacency matrix \(\mtxb{C}\).

where the weights used by node \(i\) to combine values from neighbor nodes \(j \in \Nset_i\) are given at the \(i\)-th row. The distributed averaging approach is an iterative approach, however here, we will try to find a way to limit the number of iterations necessary. This goal comes from the fact that the iterations require data from neighboring nodes leading to additional transmissions (per frame).

% We take avantage of the fact that the squared norm of the stacked solution vector can be separated into
% \begin{equation}
%   \| \h \|^2 = \sum_{j \in \Mset} \|\h_j\|^2.
% \end{equation}
We define the approximation of the global norm \(\hat{\phi}_i \approx \|\h\|^2\) at each node \(i\) at frame \(m\) as follows:
\begin{equation}
  \phi_i^{m,k+1} = \sum_{j \in \Nset_i} W_{ij} \phi_j^{m,k}\quad\text{for }k=1,\ldots,K\label{eq:dist_norm_est_iter}
\end{equation}
with the initial value at \(m=1,k=1\) defined as \(\phi_i^{1,1}=\|\h_i^{1}\|^2\).
\eqref{eq:dist_norm_est_iter} is applied \(K\) times.
We now define the frame estimate of the global squared norm as
\begin{equation}
  \hat{\phi}_i^{m+1} = \gamma_i \bar{\phi}_i^{m} + (1-\gamma) \phi_i^{m,K}
\end{equation}
where \(\bar{\phi}_i^{m} = \sum_{j \in \Nset_i} W_{ij} \|\h_i^m\|^2\) is an instantaneous weighted neighborhood average of node-wise squared norm values and \(\gamma_i\) is a mixing factor.
This factor can be chosen as a fixed value, where simulations have shown that values around \(\gamma_i = 0.02\) lead to satisfying results (cf. FIGURE) or adaptively.
For this, we use the instantaneous values to track stationarity of the squared norms by setting the factor \(\gamma_i^{m}\) proportional to the absolute difference between instantaneous neighborhood norm values of subsequent frames
\begin{equation}
  \gamma_i^{m} = \min \left\lbrace \frac{| \bar{\phi}_i^{m} - \bar{\phi}_i^{m-1} |}{\bar{\phi}_i^{m-1}},\,1\right\rbrace.
\end{equation}

The special case of \(K=1\) leads to the update step:
\begin{equation}
  \hat{\phi}_i^{m+1} = \gamma_i \sum_{j \in \Nset_i} W_{ij} \| \bar{\h}_j^{m+1} \|^2 + (1-\gamma_i) \sum_{j \in \Nset_i} w_{ij} \hat{\phi}_j^m
\end{equation}

@@

% \begin{align}
%   \w_i^{m+1} &= \operatorname{argmin} \mathcal{L}_\rho(\w,\h,\uu)\\
%   \bar{\h}_i^{m+1} &= \bar{\w}_i^{m+1} + \frac{1}{\rho} \bar{\uu}_i^m\\
%   \phi_i^1 &= \hat{\phi}_i^{m}\\
%   \phi_i^{k+1} &= \sum_{j \in \Nset_i} w_{ij} \phi_j^{k} \qquad\text{for } k=1,\ldots,K\\
%   \hat{\phi}_i^{m+1} &= \gamma \| \bar{\h}_j^{m+1} \|^2 + (1-\gamma) \phi_i^K\\
%   \h_i^{m+1} &= \frac{\bar{\h}_i^{m+1}}{\sqrt{\hat{\phi}_i^K\,M}}\\
%   \uu_i^{m+1} &= \uu_i^{m} + \rho \left(\w_i^{m+1} - \tilde{\h}_i^{m+1}\right).
% \end{align}
